Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-09-15 Final_Manuscript.Rmd
Some edits here and there

@robertsfredrickson Objectification Theory, suggests that, in addition to being steeped in a culture that objectifies women, women are objectified in actual interpersonal encounters. The negative effects of this interpersonal objectification are theorized be the strongest when it is a man doing the objectifying [@gervais2011you], but what are the consequences when a woman objectifies another woman in an interaction? Psychological researchers studying the sexual objectification of women, have in recent years explored the interpersonal process of the objectification, finding evidence for self-objectification in the target as a proximal consequence [@garcia2016objectification, @strelan2018birds, @riemer2020self]. For example, in a face-to-face mixed-gender dyadic study, @garcia2016objectification found that men’s reported objectification of their female interaction partner was associated with increased self-objectification reported by their partner. They did not include same-gender, woman-woman interacting pairs. @gervais2020social have reviewed the literature on interpersonal self-objectification and has organized the process in a theoretical model called the Social Interaction Model of Objectification (SIMO), but this model is currently only relevant for mixed-gender encounters. Although there is ample evidence that women objectify other women [@loughnan2015exploring; @puvia2013being], there are currently no studies investigating the process of interpersonal objectification in same-gender interactions. Does objectification by a female interaction partner have the same downstream negative consequences as being objectified by a male interaction partner? The current dyadic study addresses this omission using interacting pairs where both members identify as women.

##Self-Objectification
Self-objectification is a psychological process that provides a cognitive mechanism to translate the experiences of interpersonal objectification [@loughnan2015exploring] to negative mental health outcomes (e.g., anxiety, self-esteem, and cognitive performance) [@calogero2011; @moradi2008]. @calogero2011 proposes that self-objectification can be conceptualized as a learned trait, or *trait* self-objectification (TSO). Furthermore, it can also be elicited momentarily, through the media, for example, with sexualized images in movies and magazines [@morry2001magazine], or when wearing sexualizing clothing [@fredrickson1998swimsuit]. This momentary self-objectification is referred to as *state* self-objectification [@bernard2018bodies; @calogero2011; @moradi2008]. Being objectified by another person may interact TSO to effect experiences of feeling like a body, or state self-objectification [SSO; @garcia2016objectification].

##Women Objectifying Women
@harsey2020men found that self-objectification was related to the objectification of women to a similar degree for men and women. Additionally, @strelan2005women found that the more women self-objectify, the more they objectify other women. Could the process work in the other direction as well, that is, could objectification by a woman produce self-objectification, in a similar manner to mixed-gender interactions?

Past research has focused on men’s objectifying behaviors, with little attention paid to the role that women may also play in objectifying other women. @hill2008examining assessed women’s experiences of objectification from men independently from women’s experiences of objectification from other women. They found that women may be socialized not only to see themselves as objects, but perhaps to see other women as objects as well. This process, however, was not comfirmed to occur during an actual interpersonal interaction. @loughnan2015exploring found that women objectify other women to a greater extent than they objectify men. Women have also been found to objectify (dehumanized) sexualized targets presented as images [@puvia2013being], but very little is known about interpersonal sexual objectification among women and what the proximal effects of this objectification might be. We do know that objectification has more adverse consequences for women than men [@gervais2011you; @moradi2008; @saguyetal2010], however, if women objectify other women, perhaps even to a stronger degree than they objectify men, this woman to woman objectification might have detrimental outcomes.

##Interpersonal Objectification
Studies have shown that within social encounters women are gazed at more than men [@briton1995], often times feel ‘looked at’ within interpersonal interactions [@argyle1969; @miles2015stop], and are likely to internalize the objectifying gaze on their physical self [@puvia2013being; @young1979]. Moreover, perhaps the most adverse effect is that being objectified socializes girls and women to treat themselves as objects to be looked at and evaluated [@Bartky; @berger1972; @fredrickson1998swimsuit]. @gervais2020social recently reviewed the research on interpersonal objectification and organized our current theoretical understanding of this process in the Social Interaction Model of Objectification (SIMO). What is clear from @gervais2020social's review is that we know little about women-on-women interpersonal objectification. Could there also be negative consequences of this type of interpersonal objectification?

@@

There is evidence that the experience of state self-objectification in mixed-gender contexts (stranger and romantic) has negative consequences for women, but what about in the context of interactions with other women? Is the self-objectification experienced in mixed-gender interactions associated with the same negative process as that experienced in same-gender interactions? On the one hand, the ample research demonstrating that the male gaze has a particularly detrimental effect would suggest no [@saguyetal2010, @calogero2004test; @gervais2011you; @gay2010my; @yilmaz2019whose], but on the other hand, women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men]. When they do it, does it lead to self-objectification in the same way men's objectification of women does [@garcia2016objectification]? If so, does the self-objectification experienced in these same-gender interactions have the same negative consequences for authenticity in that interaction?

##Interpersonal Objectification and Authenticity
There is some evidence that reduced authenticity is a consequence of self-objectification [@garcia2016objectification; @teran2020relational]. This link has been justified by the literature on stigmatized-stigmatizer interactions [@hebl2005promoting; @richeson2003prejudice; @shelton2005expecting], viewing the experience of being objectified in an interaction as an identity threat situation [@miles2015stop; @nadal2012effects]. Further, empirical evidence reveals that objectification manifests through inauthenticity in romantic relationships [@brunelletal2010], adverse attitudes in regard to career aspirations, and a decrease in concentration and impairment in female cognitive performance [@kahalon2018don; @quinnetal]. When a woman is objectified by a man, and subsequently experiences self-objectification, the intergroup nature of this encounter might trigger identity threat. Thus, perhaps there are fewer negative consequences when a woman is objectified by another woman, specifically, woman-woman interpersonal objectification processes might diverge from mix-gender interpersonal objectification processes precisely because they are not intergroup interactions (at least with respect to gender identity). A woman objectified by another woman may not be having the same negative consequences that cascade from situations that trigger group-based identity threat [@deaux1987putting; @dovidio2006nonverbal; @hebl2005promoting], but feeling like a body, rather than a full human, in any interaction (intergroup or ingroup) may be enough to reduce women's feelings of authenticity and social competence, regardless of the gender of the objectifyer [@tolman2006looking; @teran2020relational].

@@

Data from two demographically equivalent higher education institutions in the Northeast United States were combined to create the final analysis sample (*N =* `r nrow(clarksmith_women)`) used in this study. In the measures section that follows we refer to them as Sample 1 and Sample 2. Thirty-two previously unacquainted self-identifying female dyads (64 total participants) from institutions in the Northeast of the United States participated in this study. Sample 1 (*N =* `r ns[3]`) is from a co-ed liberal arts college and Sample 2 (*N =* `r ns[4]`) is from a women’s liberal arts college. More specifically, twelve of the pairs were students at a co-ed liberal arts college, while the remaining twenty pairs attended a women’s liberal arts college. [Footnote: Initially, data was collected from same-sex and mixed-sex dyads that comprised of male and female gendered individuals. Sample 1 originally consisted of twenty-two pairs, twelve men and thirty-two women. Twenty-three pairs made up of forty-three women and one man, as well as two participants who did not identify with either gender category, formed Sample 2. For consistency, we limited participant data to same sex female pairs at the two colleges.] All results presented below are from models including sample as a control variable.

@@

#Discussion
The current study tested whether the model of interpersonal objectification and state self-objectification (SSO) used in @garcia2016objectification replicates in a same-gender dyadic sample of women only. Although past research has found that women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men], this is the first study to test if *interpersonal* other-objectification by women is related to state self-objectification in their woman-identified interaction partners. As hypothesized, the current study did find a significant relationship between a woman's interaction partner's report of having objectified her and her own post-interaction feelings of self-objectification. That is, there was a significant partner effect of other-objectification on SSO. This effect extends the equivalent relationship found in mixed-gender interactions to the context of same-gender interactions between women. Thus, evidence suggests that it is not only the male gaze, and male other-objectification that is related to women's state self-objectification, but being objectified by another woman can also result in SSO, at least in the context of a scenario where they know they are being evaluated as a potential dating partner. [Asha:Or … thus evidence provides support for the claim that women being objectified by another woman result in sso in ways that are not uniquely dissimilar from that of xyz]

As is the case in all correlational studies, we cannot be sure about the causal direction. It could be that women's SSO causes them to be objectified by their interaction partner. This interpretation is theoretically interesting given that, in both the @garcia2016objectification study and in the current study, it is *women's* SSO that relates to being objectified by one's partner. However, Objectification Theory [@robertsfredrickson], as well as past experimental studies [@saguyetal2010; CITE], suggest that the causal flow is from other-objectification to SSO. Although, the current study found evidence for the partner effect of other-objectification on SSO, there was a lack of evidence for the connections between SSO and downstream negative consequences.

Where the results of the current study diverge most notably from the results of studies testing interpersonal objectification among mixed-gender dyads is the lack of evidence for relationships between SSO and interaction inauthenticity. In addition, we found no evidence of *direct* or *indirect* relationships between SSO and cognitive functioning, relationship agency, and career aspirations. Nor was there evidence of *direct* or *indirect* relationships between other-objectification and any of these outcomes. This is somewhat suprising given the extant evidence linking SSO and cognitive functioning [@fredrickson1998swimsuit; @quinn2011performance; and see @moradi2008 for a review] and the small, but extant research on interpersonal other-ojectification and cognitive functioning [@logel2009interacting; @garcia2016objectification]. This lack of evidence could potentially signal diverging processes between women's experiences with interpersonal objectification from men and interpersonal objectification from women. There is quite a bit of evidence suggesting that the male gaze is particularly detrimental [@calogero2004test; @robertsfredrickson; @gervais2011you; @gervais2013my], and perhaps the self-objectification experienced within an interaction with a women is not as harmful as the self-objectification experienced within an interaction with a man. As a strong note of caution, it is best not to interpret a null result as evidence of no relationship.

@@

Although we did not find a connection between SSO and authenticity, we did find significant positive relationships between authenticity and relationship agency, and authenticity and career aspirations. The relationship between authenticity and cognitive functioning was also estimated as positive, but was not statistically signficant. Again, due to the lack of connection between SSO and authenticity, we found no evidence of *indirect* relationships between SSO and these outcome variables. This evidence of a relationship between authenticity and the relational outcome variables (i.e., relationship agency and career aspirations) provides evidence that corroborates past findings that felt authenticity in interactions is important for healthy relationship functioning [@garcia2016objectification; @teran2020relational]. Just as authenticity has been found to be important in intergroup interactions [@brunelletal2010; @garcia2016objectification; @tolman2006looking], we again find more evidence here that disruptions in feelings of authenticity can negatively impact relationships beyond the current partner. Although the current study did not find a connection between authenticity and SSO, this seems theoretically to be a natural connection, and future work must explore when and how SSO leads to inauthenticity in interpersonal objectification situations.

@@

>>>>

@robertsfredrickson Objectification Theory, suggests that, in addition to being steeped in a culture that objectifies women, women are objectified in actual interpersonal encounters. The negative effects of this interpersonal objectification for women are theorized be the strongest when it is a man doing the objectifying [@calogero2004test; @gervais2011you; @gay2010my], but what are the consequences when a woman objectifies another woman in an interaction? Psychological researchers studying the sexual objectification of women, have in recent years explored the interpersonal process of the objectification, finding evidence for self-objectification in the target as a proximal consequence of being objectified [@garcia2016objectification, @strelan2018birds, @riemer2020self]. In a recent review, @gervais2020social organized the extant literature on interpersonal self-objectification and proposed a theoretical model called the Social Interaction Model of Objectification (SIMO), however this model focuses on mixed-gender encounters. Although there is evidence that women objectify other women [@loughnan2015exploring; @puvia2013being], there are currently no studies investigating the process of interpersonal objectification in interactions between women. Is objectification by a female interaction partner related to an increase in self-objectification for the woman being objectified? Further, does objectification by a woman have the same downstream negative consequences for women as being objectified by a male interaction partner? The current study begins to address these questions by investigating interpersonal objectification among interacting pairs where both partners identify as women.

##Self-Objectification
Self-objectification is a psychological process that provides a cognitive mechanism to translate the experiences of interpersonal objectification [@loughnan2015exploring] to negative mental health outcomes (e.g., anxiety, self-esteem, and cognitive performance) [@calogero2011; @moradi2008]. @calogero2011 proposes that self-objectification can be conceptualized as a learned trait, or *trait* self-objectification (TSO). Furthermore, self-objectification can also be elicited momentarily, for example, when viewing sexualized images in movies and magazines [@morry2001magazine], when trying on sexualizing clothing [@fredrickson1998swimsuit], or during video chat when you believe your interaction partner can only see your body [not face; @saguyetal2010]. This momentary self-objectification is referred to as *state* self-objectification [SSO; @bernard2018bodies; @calogero2011; @moradi2008] and is characterized by feeling like a body rather than a full self within a particular moment, instance, or context [@robertsfredrickson].

There is evidence that being objectified by another person during an interaction elicits SSO. For example, in a face-to-face mixed-gender dyadic study, @garcia2016objectification found that men’s reported objectification of their female interaction partner was associated with increased self-objectification as reported by their female partner. Further, this negative effect of the men's objectification on women's SSO was strongest for women higher in TSO. The current study uses this same dyadic paradigm to study same-gender, woman-woman interacting pairs.

##Interpersonal Objectification
Studies have shown that within social encounters women are gazed at more than men [@briton1995], often times feel ‘looked at’ within interpersonal interactions [@argyle1969; @miles2015stop], and are likely to internalize the objectifying gaze on their physical self [@puvia2013being; @young1979]. Interpersonal objectification has been found to have negative consequences for women. @gervais2011you found that an objectifying gaze by a male interaction partner (confederate) was associated with lower math performance for women, than being objectified by a female interaction partner. Moreover, perhaps the most adverse effect of interpersonal objectificaiton is that being objectified socializes girls and women to treat themselves as objects to be looked at and evaluated [@Bartky; @berger1972; @fredrickson1998swimsuit]. @gervais2020social recently reviewed the research on interpersonal objectification and organized our current theoretical understanding of this process in the Social Interaction Model of Objectification (SIMO). What is clear from @gervais2020social's review is that we know little about women-on-women interpersonal objectification. Could there also be negative consequences of this type of interpersonal objectification?

@@

There is evidence that the experience of state self-objectification in mixed-gender contexts (stranger and romantic) has negative consequences for women, but what about in the context of interactions with other women? Is the self-objectification experienced in mixed-gender interactions associated with the same negative process as that experienced in same-gender interactions? The ample research demonstrating that the male gaze has a particularly detrimental effect for women would suggest no [@saguyetal2010, @calogero2004test; @gervais2011you; @gay2010my; @yilmaz2019whose], but, there is evidence that women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men] and to would be helpful to understand the consequencs of this intragroup objectification for women. When women objectify other women, does it lead to self-objectification in the same way men's objectification of women does [@garcia2016objectification]? If so, does the self-objectification experienced in these same-gender interactions have the same negative consequences for authenticity in that interaction?

There is some evidence that reduced authenticity is a consequence of self-objectification [@garcia2016objectification; @teran2020relational]. This link has been justified by the literature on stigmatized-stigmatizer interactions [@hebl2005promoting; @richeson2003prejudice; @shelton2005expecting], viewing the experience of being objectified in an interaction as an identity threat situation [@miles2015stop; @nadal2012effects]. Further, empirical evidence reveals that objectification manifests through inauthenticity in romantic relationships [@brunelletal2010], adverse attitudes in regard to career aspirations, and a decrease in concentration and impairment in female cognitive performance [@kahalon2018don; @quinnetal]. When a woman is objectified by a man, and subsequently experiences self-objectification, the intergroup nature of this encounter might trigger identity threat. Thus, perhaps there are fewer negative consequences when a woman is objectified by another woman, specifically, woman-woman interpersonal objectification processes might diverge from mix-gender interpersonal objectification processes precisely because they are not intergroup interactions (at least with respect to gender identity). A woman objectified by another woman may not be having the same negative consequences that cascade from situations that trigger group-based identity threat [@deaux1987putting; @dovidio2006nonverbal; @hebl2005promoting], but feeling like a body, rather than a full human, in any interaction (intergroup or ingroup) may be enough to reduce women's feelings of authenticity and social competence, regardless of the gender of the objectifyer [@tolman2006looking; @teran2020relational].

##Women Objectifying Women
Past research investigating women objectifying other women has focused on the psychological conditions that led women to objectify women. @harsey2020men found that self-objectification was related to the objectification of women to a similar degree for male and female participants. Along the same lines, @strelan2005women found that the more women self-objectify, the more they objectify other women. Could the process work in the other direction as well, that is, could objectification by a woman produce self-objectification, in a similar manner to objectification by a man? @hill2008examining assessed women’s experiences of objectification from men independently from women’s experiences of objectification from other women. They found that women may be socialized not only to see themselves as objects, but perhaps to see other women as objects as well. This process, however, was not comfirmed to occur during an actual interpersonal interaction. @loughnan2015exploring found that women objectify other women to a greater extent than they objectify men. Women have also been found to objectify (dehumanized) sexualized targets presented as images [@puvia2013being], but very little is known about interpersonal sexual objectification among women and what the proximal effects of this objectification might be. We do know that objectification has more adverse consequences for women than men [@gervais2011you; @moradi2008; @saguyetal2010], however, if women objectify other women, perhaps even to a stronger degree than they objectify men, this woman to woman objectification might have detrimental outcomes.

@@

Data from two demographically similar higher education institutions in the Northeast United States were combined to create the final analysis sample (*N =* `r nrow(clarksmith_women)`) used in this study. In the measures section that follows we refer to them as Sample 1 and Sample 2. Thirty-two previously unacquainted self-identifying female dyads (64 total participants) from institutions in the Northeast of the United States participated in this study. Sample 1 (*N =* `r ns[3]`) is from a co-ed liberal arts college and Sample 2 (*N =* `r ns[4]`) is from a women’s liberal arts college. More specifically, twelve of the pairs were students at a co-ed liberal arts college, while the remaining twenty pairs attended a women’s liberal arts college. [Footnote: Initially, data was collected from same-sex and mixed-sex dyads that comprised of male and female gendered individuals. Sample 1 originally consisted of twenty-two pairs, twelve men and thirty-two women. Twenty-three pairs made up of forty-three women and one man, as well as two participants who did not identify with either gender category, formed Sample 2. For consistency, we limited participant data to same sex female pairs at the two colleges.] All results presented below are from models including sample as a control variable.

@@

#Discussion
The current study tested whether the model of interpersonal objectification and state self-objectification (SSO) used in @garcia2016objectification replicates in a same-gender dyadic sample of women only. Although past research has found that women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men], this is the first study to test if *interpersonal* other-objectification by women is related to state self-objectification in their woman-identified interaction partners. As hypothesized, the current study did find a significant relationship between a woman's interaction partner's report of having objectified her and her own post-interaction feelings of self-objectification. That is, there was a significant partner effect of other-objectification on SSO. This effect extends the equivalent relationship found in mixed-gender interactions to the context of same-gender interactions between women. Thus, evidence suggests that it is not only the male gaze, and male other-objectification that is related to women's state self-objectification, but being objectified by another woman can also result in SSO, at least in the context of a scenario where they know they are being evaluated as a potential dating partner.

MORE ABOUT THIS PARTNER EFFECT

As is the case in all correlational studies, we cannot be sure about the causal direction. It could be that women's SSO causes them to be objectified by their interaction partner. This interpretation is theoretically interesting given that, in both the @garcia2016objectification study and in the current study, it is *women's* state self-objectification that relates to being objectified by one's partner. However, Objectification Theory [@robertsfredrickson], as well as past experimental studies [@saguyetal2010], suggest that the causal flow is from other-objectification to SSO.

Where the results of the current study diverge most notably from the results of studies testing interpersonal objectification among mixed-gender dyads is the lack of evidence for relationships between SSO and interaction inauthenticity. In addition, we found no evidence of *direct* or *indirect* relationships between SSO and cognitive functioning, relationship agency, and career aspirations. Nor was there evidence of *direct* or *indirect* relationships between other-objectification and any of these outcomes. This is somewhat suprising given the extant evidence linking SSO and cognitive functioning [@fredrickson1998swimsuit; @quinn2011performance; and see @moradi2008 for a review] and the research on interpersonal other-ojectification and cognitive functioning [@logel2009interacting; @garcia2016objectification; @gervais2011you]. This lack of evidence could potentially signal diverging processes between women's experiences with interpersonal objectification from men and interpersonal objectification from women. There is quite a bit of evidence suggesting that the male gaze is particularly detrimental [@calogero2004test; @robertsfredrickson; @gervais2011you; @gervais2013my], and perhaps the self-objectification experienced within an interaction with a women is not as harmful as the self-objectification experienced within an interaction with a man. However, as a strong note of caution, it is best not to interpret a null result as evidence of no relationship.

@@

Although the urrent study did not find a connection between SSO and authenticity, we did find significant positive relationships between authenticity and relationship agency, and authenticity and career aspirations. The relationship between authenticity and cognitive functioning was also estimated as positive, but was not statistically signficant. Again, due to the lack of connection between SSO and authenticity, we found no evidence of *indirect* relationships between SSO and these outcome variables. This evidence of a relationship between authenticity and the relational outcome variables (i.e., relationship agency and career aspirations) provides evidence that corroborates past findings that felt authenticity in interactions is important for healthy relationship functioning [@garcia2016objectification; @teran2020relational]. Just as authenticity has been found to be important in intergroup interactions [@brunelletal2010; @garcia2016objectification; @tolman2006looking], we again find more evidence here that disruptions in feelings of authenticity can negatively impact relationships beyond the current partner. Although the current study did not find a connection between authenticity and SSO, this seems theoretically to be a natural connection, and future work must explore when and how SSO leads to inauthenticity in interpersonal objectification situations.

@@

Men-men
